## Comparison between Aranet4 and SenseAir S8 CO~2~ sensors

We gathered data over the course of a single day to compare the CO~2~
measurements from an [Aranet4](https://aranet.com/products/aranet4/)
and two different [SenseAir S8
LP](https://senseair.com/products/size-counts/s8-lp/) CO~2~ sensors,
one controlled by an Arduino Nano and the other controlled by an
Arduino Nano Every.

The monitors were in a common area, sitting all together. We had
closed all doors and windows in the morning, and then opening them again
late afternoon.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev="svg",
                      message=FALSE, warning=FALSE,
                      fig.width=11, fig.height=6)
options(scipen=10, width=110)
library(chron)   # for times
library(broman)  # for graphs
```

```{r load_data, include=FALSE}
x <- read.csv("data.csv", comment.char="#")

x$time <- chron(times=paste0(x$time, ":00"))
```

Here's a plot of the three sensors' values over time. The Aranet4 is
in black; the two SenseAir S8s are in blue and pink.

```{r plot_data}
par(mar=c(3.1,4.1,0.6,0.6))
yl <- range(x[,2:4])
pink <- "violetred"
blue <- "slateblue"

xlab <- c("6am", "9am", "noon", "3pm", "6pm", "9pm")
xticks <- chron(times=paste0(c("06", "09", "12", "15", "18", "21"), ":00:00"))

ylab <- expression(paste(CO[2], " (in ppm)"))
devices <- c("Aranet4", paste0("SenseAir S8 #", 1:2))

grayplot(x[,1], x[,2], type="l", lwd=2, mgp.y=c(3.1,0.4,0.4),
         xlab="Time", ylab=ylab,
         ylim=range(yl), xat=NA, vlines=xticks)
axis(side=1, at=xticks, label=xlab, tick=FALSE, mgp=c(0,0.3,0.03))
lines(x[,1], x[,3], col=pink, lwd=2)
lines(x[,1], x[,4], col=blue, lwd=2)
legend("topright", lwd=2, col=c("black", blue, pink),
       devices, bg="gray90")
```

It had started raining, so we closed up the house at 8:30am or so, and
the CO~2~ gradually increased, reaching a peak of about 1200 ppm at
about 5pm. At that point, we opened windows and doors again, and the
CO~2~ quickly plunged back to an acceptable level (in less than an hour).

The three sensors give qualitatively similar information, though one
of the SenseAir S8 sensors is noticeably shifted down, and gave a peak
reading of `r max(x[,3])` ppm, vs the peak of `r max(x[,2])` ppm from
the Aranet4.

Directly plotting the differences, taking the Aranet4 values as the standard,
makes those differences more clear.

```{r plot_differences}
par(mar=c(3.1,4.1,0.6,0.6))
yl <- c(-100,100)

dif_ylab <- expression(paste("diff in", CO[2], " (in ppm)"))

grayplot(x[,1], x[,4]-x[,2], col=blue, type="l", lwd=2,
         xlab="Time", ylab=dif_ylab,
         ylim=yl, xat=NA, vlines=xticks,
         hlines=c(-100, -50,0,50,100), yaxs="i",
         hlines.col=c("white", "orange")[c(1,1,2,1,1)])
axis(side=1, at=xticks, label=xlab, tick=FALSE, mgp=c(0,0.3,0.03))
lines(x[,1], x[,3]-x[,2], col=pink, lwd=2)
legend("topright", lwd=2, col=c(blue, pink),
       devices[-1], bg="gray90")
```

One of the SenseAir S8 sensors tends to be above the Aranet4 by
an average of `r round(mean(x[,4])-mean(x[,2]))` ppm, while the
other tends to be below by `r -round(mean(x[,3])-mean(x[,2]))` ppm.
The shifts are relatively constant, though the first sensor tends to be off
a bit more when then readings are low, and the second sensor tends to
be off a bit more when the readings are high.

If we take the Aranet4 readings to be the gold standard, the RMS error in
the two SenseAir S8 sensors are
`r round(sqrt(mean((x[,4]-x[,2])^2)))` and
`r round(sqrt(mean((x[,3]-x[,2])^2)))` ppm, respectively.
The largest discrepancies were
`r round(max(abs(x[,4]-x[,2])))` and `r round(max(abs(x[,3]-x[,2])))`
ppm, respectively.

If we adjust for the overall shifts in the values, the RMS errors in
the two SenseAir S8 sensors are reduced to
`r round(sd(x[,4]-x[,2]))` and
`r round(sd(x[,3]-x[,2]))` ppm, respectively.
(These are just the regular SDs of the errors.)

In particular, most of the error in the second sensor is do to a shift that might
be corrected by calibration.

**In conclusion**, based on this admittedly limited data (from a single
day), these two SenseAir S8 sensors seem to provide equivalent
information as the much more expensive Aranet4.

---

The source for this document, including the data, is at
<https://github.com/karlduino/CO2monitor/tree/main/docs/Comparison>

Details on the design and assembly of the Arduino-based CO~2~ monitor are at
<https://karlduino.github.io/CO2monitor>.
