## Comparison between Aranet4 and SenseAir S8 CO2 sensors

We gathered data over the course of a single day to compare the CO2
measurements from an [Aranet4](https://aranet.com/products/aranet4/)
and two different [SenseAir S8
LP](https://senseair.com/products/size-counts/s8-lp/) CO2 sensors. The
monitors were in a common area, sitting all together. We had closed
all doors and windows mid-morning, and then opening them again late
afternoon.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev="svg",
                      message=FALSE, warning=FALSE,
                      fig.width=11, fig.height=6)
options(scipen=10, width=110)
library(chron)   # for times
library(broman)  # for graphs
```

```{r load_data, include=FALSE}
x <- read.csv("data.csv", comment.char="#")

x$time <- chron(times=paste0(x$time, ":00"))

# calibrate the S8s vs the Aranet4 by subtracting their averages
x[,5:6] <- t(t(x[,3:4])-colMeans(x[,3:4])+mean(x[,2]))
```

Here's a plot of the three sensors' values over time. The Aranet4 is
in black; the two Sensair S8s are in blue and pink.

```{r plot_data}
par(mar=c(3.1,4.1,0.6,0.6))
yl <- range(x[,2:4])
pink <- "violetred"
blue <- "slateblue"

xlab <- c("6am", "9am", "noon", "3pm", "6pm", "9pm")
xticks <- chron(times=paste0(c("06", "09", "12", "15", "18", "21"), ":00:00"))

ylab <- expression(paste(CO[2], " (in ppm)"))
devices <- c("Aranet4", paste0("SenseAir S8 #", 1:2))

grayplot(x[,1], x[,2], type="l", lwd=2, mgp.y=c(3.1,0.4,0.4),
         xlab="Time", ylab=ylab,
         ylim=range(yl), xat=NA, vlines=xticks)
axis(side=1, at=xticks, label=xlab, tick=FALSE, mgp=c(0,0.3,0.03))
lines(x[,1], x[,3], col=pink, lwd=2)
lines(x[,1], x[,4], col=blue, lwd=2)
legend("topright", lwd=2, col=c("black", blue, pink),
       devices, bg="gray90")
```
